---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# omock

<!-- badges: start -->
[![R-CMD-check](https://github.com/oxford-pharmacoepi/omock/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/oxford-pharmacoepi/omock/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

## Installation

You can install the development version of omock using:

``` r
# install.packages("devtools")
devtools::install_github("oxford-pharmacoepi/omock")
```

## Example

With omock we can quickly make a simple mock of OMOP CDM data.

```{r, warning=FALSE, message=FALSE}
library(omopgenerics)
library(omock)
library(dplyr)
```

We first start by making an empty cdm reference. This includes the person and observation tables (as they are required) but they are currently empty.
```{r}
cdm <- emptyCdmReference(cdmName = "mock")
cdm$person %>% 
  glimpse()
cdm$observation_period %>% 
  glimpse()
```

Once we have have our empty cdm reference, we can quickly add a person table with a specific number of individuals.
```{r}
cdm <- cdm %>% 
  omock::mockPerson(nPerson = 1000)

cdm$person %>% 
  glimpse()
```

We can then fill in the observation period table for these individuals.
```{r}
cdm <- cdm %>% 
  omock::mockObservationPeriod()

cdm$observation_period %>% 
  glimpse()
```
