---
title: "Building a bespoke mock cdm"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{a04_Building_a_bespoke_mock_cdm}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup,message=FALSE, warning=FALSE}
library(omock)
library(dplyr)
library(lubridate)
```

The mockCdmFromTable() function is designed to create a mock Common Data Model (CDM) object from user-defined tables. This function facilitates the generation of a synthetic dataset for testing and development purposes.

Function Parameters

- cdm: The name of the CDM object to be modify.

- cohortTable: A list of user-defined cohort tables. Default is NULL.

- seed: A random seed for reproducibility. Default is 1.

Return Value

The function returns an OMOP CDM object with the user-defined cohort tables, including the addition of observation period and person tables that matches the user-defined cohort tables to make it a valid OMOP CDM object.

For example if you want to create a CDM object based on belows bespoke cohorts. You can do it simple using the mockCdmFromTable() functions.

```{r, warning = FALSE}
# Define a list of user-defined cohort tables
cohortTables <- list(
  cohort1 = tibble(
    subject_id = 1:10,
    cohort_definition_id = rep(1, 10),
    cohort_start_date = as.Date('2020-01-01') + 1:10,
    cohort_end_date = as.Date('2020-01-01') + 11:20
  ),
  cohort2 = tibble(
    subject_id = 11:20,
    cohort_definition_id = rep(2, 10),
    cohort_start_date = as.Date('2020-02-01') + 1:10,
    cohort_end_date = as.Date('2020-02-01') + 11:20
  )
)

# Create a mock CDM object from the user-defined tables
cdm <- mockCdmReference() |> mockCdmFromTable(cohortTable = cohortTables)

cdm |> glimpse()

```

