---
title: "synthetic_datasets"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{synthetic_datasets}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r}
library(omock)
```

# Introduction

As seen in other vignettes **omock** provides you with functionality to build. synthetic datasets. **omock** also provides some prebuilt synthetics datasets, those datasets are widely available and were created by the OHDSI community.

## Avialable datasets

The available datasets are listed below:

```{r, echo=FALSE}
niceSize <- function(x) {
  purrr::map_chr(x, \(x) {
    if (x < 1024) {
      x <- fprintf("%i B", x)
    } else if (x < 1024**2) {
      x <- x / 1024
      x <- paste(formatC(x, digits=3, format="fg", flag="#"), "kB")
    } else if (x < 1024**3) {
      x <- x / 1024**2
      x <- paste(formatC(x, digits=3, format="fg", flag="#"), "MB")
    } else {
      x <- x / 1024**3
      x <- paste(formatC(x, digits=3, format="fg", flag="#"), "GB")
    }
    stringr::str_squish(x)
  })
}
mockDatasets |>
  dplyr::mutate(
    size = niceSize(.data$size),
    link = paste0("[🔗](", .data$url, ")"),
    dplyr::across(
      c("number_individuals", "number_records", "number_concepts"), 
      \(x) format(x, big.mark = ",")
    )
  ) |>
  dplyr::select(
    "datasetName" = "dataset_name", "CDM name" = "cdm_name", 
    "CDM version" = "cdm_version", "Size" = "size", 
    "Number individuals" = "number_individuals",
    "Number records" = "number_records", "Number concepts" = "number_concepts", 
    "Link" = "link"
  ) |>
  gt::gt() |>
  gt::fmt_markdown("Link")
```

For more details on those synthetic datasets you can check the [OmopSketch](https://ohdsi.github.io/OmopSketch/) ShinyApp: <https://dpa-pde-oxford.shinyapps.io/OmopSketchCharacterisation/> that characterise those datasets.

You can also check programatically which are the synthetic datasets that you can use with:

```{r}
availableMockDatasets()
```

## Download a dataset

To prevent having to download the dataset everytime that you want to set up a reference to it, it is recommended to set up a permanent folder where the synthetic datasets are stored, this allows the user to have to download each dataset only once. To set up a permanent location for your dataset please create an environmental variable (`usethis::edit_r_environ()`) pointing to an existing folder:

```
OMOP_DATA_FOLDER="path/to/my/folder"
```

This folder is in fact defined by [omopgenerics](https://darwin-eu.github.io/omopgenerics/) and it is used also by other packages. You can check the folder by using the following function:

```{r}
omopDataFolder()
```

If you would have set up an environment variable the message of temporary folder would not appear and you would see the path to you folder.

You can download a dataset using ``:

```{r}
downloadMockDataset(datasetName = "synthea-covid19-10k")
```

This will download the dataset and store it as a zip file in you `OMOP_DATA_FOLDER`:

```{r}
list.files(path = omopDataFolder(), recursive = TRUE)
```

Note this step is not needed as each dataset will be downloaded (if not stored already) when you create a cdm reference with it.

##

